---
title: "HW2_nh2706"
author: "Niklas Hess"
date: 2022-10-02
output: github_document
---

```{r, message = FALSE, echo = FALSE}
library(tidyverse)
```

# Problem 1

**Omitted answer, as solutions have already been provided**

# Problem 2

# Problem 3

## 3.1) pols-month.csv related tasks

```{r}
# Loading the data and cleaning the variables
polsm_df = read_csv("./Data/pols-month.csv", show_col_types = FALSE) %>%
  janitor::clean_names() %>%

# Separating variable MON into individual pieces
  separate(mon, into = c("year", "month", "day")) %>%

# Arranging entire dataset based on the year and month
  arrange(year, as.integer(month)) %>%
  
# Replacing month number with month name and creating a president variable taking values gop and dem
  mutate(year = as.integer(year),
         month = month.name[as.integer(month)],
         president = if_else(prez_dem == 1, "dem", "gop")
         ) %>%

# Removing prez_dem, prez_gop, and day variables
  select(-day, -prez_dem, -prez_gop)
```

## 3.2) snp.csv related tasks

```{r}
# Loading the data and cleaning the variables
snp_df = read_csv("./Data/snp.csv", show_col_types = FALSE) %>%
  janitor::clean_names() %>%

# Separating variable DATE into individual pieces, and modifying years to show long version
  separate(date, into = c("month", "day", "year")) %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         year = ifelse(year >= 23, year + 1900, year + 2000)
         ) %>%

# Arranging according to year and month, organizing so that year and month are the leading columns, and converting month to long-format
  relocate(year) %>%
  arrange(year, month) %>%
  mutate(month = month.name[as.integer(month)]) %>%

# Removing day variables
  select(-day)
```

## 3.3) unemployment.csv related tasks

```{r}
# Loading the data
unemployment_df = read_csv("./Data/unemployment.csv", show_col_types = FALSE) %>%

# Switching from “wide” to “long” format
  pivot_longer("Jan":"Dec",
        names_to = "month",
        values_to = "unemployment"
        ) %>%

# Ensuring that key variables have the same name, and take the same values
  janitor::clean_names() %>%
  mutate( month = month.name[match(month, month.abb)],
          year = as.integer(year)
         )
```

## 3.4) Merging Datasets

```{r, message = FALSE}
# Merging snp into pols
alldata_df = full_join(polsm_df, snp_df)

# Merging unemployment into the result
alldata_df = full_join(alldata_df, unemployment_df)
```


## 3.5) Summary of datasets

-   The `pols-month.csv` dataset details:
    -   Dimensions:
        -   The dataset has `r nrow(polsm_df)` 
        -   The dataset has `r ncol(polsm_df)`
    -   The years within the dataset range from `r min(polsm_df$year)` to `r max(polsm_df$year)`
    -   The key variables are `r names(polsm_df)`

-   The `snp.csv` dataset details:
    -   Dimensions:
        -   The dataset has `r nrow(snp_df)` 
        -   The dataset has `r ncol(snp_df)`
    -   The years within the dataset range from `r min(snp_df$year)` to `r max(snp_df$year)`
    -   The key variables are `r names(snp_df)`

-   The `unemployment.csv` dataset details:
    -   Dimensions:
        -   The dataset has `r nrow(unemployment_df)` 
        -   The dataset has `r ncol(unemployment_df)`
    -   The years within the dataset range from `r min(unemployment_df$year)` to `r max(unemployment_df$year)`
    -   The key variables are `r names(unemployment_df)`

-   The `merged` dataset details:
    -   Dimensions:
        -   The dataset has `r nrow(alldata_df)` 
        -   The dataset has `r ncol(alldata_df)`
    -   The years within the dataset range from `r min(alldata_df$year)` to `r max(alldata_df$year)`
    -   The key variables are `r names(alldata_df)`

# END OF HW2
